Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period

### 2020-04-14 analysis/paper/paper.Rmd
edits on KWL background section based on BF's comment

Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that during 2001-2004 in advance of a water diversion project and road bridge construction. The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007]. The nearly 2 m thick archaeological deposits reveal a large amount of artifacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement. Artifact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance. Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from (Figure \@ref(fig:KWL-sam-area)). In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures. At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction.

@@

The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BP) and a lower component (700-1200 AD, 1200-800 BP) separated by a sterile deposit spanning c. 150 years. These component divisions are based on the differences in the color and texture of the deposit. The interpretation of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007]. There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data). We focus on the upper component because only this component spans the periods of pre-European contact, European presence, and Chinese presence. In the upper component, all excavation squares in our sampling area show signs of continuous human occupation during each of the three phases. Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011]. However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artifacts to the pre-European, European, and the Chinese periods.

@@

```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan. The dark line represents the summed probabilities of all radiocarbon ages, and the grey lines in the background are the probabilities of individual ages. Grey rectangles indicate the approximate chronology of the major archaeological components of the deposit. For the upper component, the blue line indicates the start of European presence, while the red line is the Chinese presence. Ages calibrated with the Bchron package (Parnell et al. 2008)"}

@@

table_kw_dates <-
  kw_dates %>% 
  mutate(`Lab code` = str_remove(`Lab code`, " "),
         `Uncalibrated Age BP` = `Date BP`) %>% 
  mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>%
  mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
  left_join(kw_age_cali_df_tidy, "Pit-Layer") %>% 
  mutate(ave = (`2.5%` + `97.5%`)/ 2) %>% 
  mutate(ave = ifelse(is.na(ave), 200, ave)) %>%  
  arrange(ave) %>% 
  mutate(`Calibrated Age BP (95% credible interval)` = 
           ifelse(is.na(`Calibrated date (95% conf. level)`), 
                  "", 
                  `Calibrated date (95% conf. level)`)) %>% 
  mutate(`Uncalibrated Age BP` = str_remove_all(`Uncalibrated Age BP`, "yr")) %>% 
  dplyr::select(`Lab code`, 
                `Pit-Layer`, 
                `Above mean sea level (cm)`, 
                `Uncalibrated Age BP`, 
                `Calibrated Age BP (95% credible interval)`, 
                Context)

kableExtra::kable(table_kw_dates,
             caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal13 Atmospheric curve. The codes in the context column refer to the excavator's feature labels, cf. Chen 2007")
```

>>>>

Information about Kiwulan (Figure \@ref(fig:KWL-map)) comes from a rescue archaeology project that was conducted between 2001-2004 in advance of a water diversion project and road bridge construction. The excavations used 2 mm and 1.5 mm mesh screens and covered eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007]. The nearly 2 m thick archaeological deposits reveal a large amount of artifacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-term settlement. Artifact locations were recorded to the 2 x 2 m sub-square they were recovered in; they lack individual point provenance. Based on the continuity of deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from (Figure \@ref(fig:KWL-sam-area)). In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures. At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction.

@@

The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BP) and a lower component (700-1200 AD, 1200-800 BP) separated by a sterile deposit spanning c. 150 years [@Chen2007]. The sterility may be related to dry weather leading to site abandonment according to pollen analysis [@Lin2015]. These component divisions are based on the differences in the color and texture of the deposit. Whether these two components represent a culture continuity from the same people is still under debate [@Chiu2004; @Chen2007]. There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data). We focus on the upper component because only this component spans the periods of pre-European contact, European presence, and Chinese presence. In the upper component, all excavation squares in our sampling area show signs of continuous human occupation during each of the three phases. Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011]. However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artifacts to the pre-European, European, and the Chinese periods.

@@

```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan. The dark line represents the summed probabilities of all radiocarbon ages, and the grey lines in the background are the probabilities of individual ages. Grey rectangles indicate the approximate chronology of the major archaeological components of the deposit. For the upper component, the red line indicates the start of European presence, while the blue line is the Chinese presence. Ages calibrated with the Bchron package (Parnell et al. 2008)"}

@@

table_kw_dates <-
  kw_dates %>% 
  mutate(`Lab code` = str_remove(`Lab code`, " "),
         `Uncalibrated Age BP` = `Date BP`) %>% 
  mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>%
  mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
  left_join(kw_age_cali_df_tidy, "Pit-Layer") %>% 
  mutate(ave = (`2.5%` + `97.5%`)/ 2) %>% 
  mutate(ave = ifelse(is.na(ave), 200, ave)) %>%  
  arrange(ave) %>% 
  mutate(`Calibrated Age BP (95% credible interval)` = 
           ifelse(is.na(`Calibrated date (95% conf. level)`), 
                  "", 
                  `Calibrated date (95% conf. level)`)) %>% 
  mutate(`Uncalibrated Age BP` = str_remove_all(`Uncalibrated Age BP`, "yr")) %>% 
  dplyr::select(`Lab code`, 
                `Pit-Layer`, 
                `Depth (cm)`, 
                `Uncalibrated Age BP`, 
                `Calibrated Age BP (95% credible interval)`, 
                Context)

kableExtra::kable(table_kw_dates,
             caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal13 Atmospheric curve. All depth is recorded in centimeters above mean sea level. The codes in the context column refer to the excavator's feature labels, cf. Chen 2007")
```
